---
title: "Intro_RStudio"
output: html_document
date: "2025-02-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "Introduction to R in RStudio Cloud"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

# Welcome to R & RStudio Cloud
R is a powerful programming language used for **data analysis, visualization, and statistics**. In this lesson, you'll learn the basics of working with R.

---

## Getting Started
When you open RStudio Cloud, you’ll see:
- **Console (bottom left)**: Runs your R commands.
- **Script Editor (top left)**: Where you write and save your code.
- **Environment & Files (top right)**: Stores variables and datasets.
- **Plots & Help (bottom right)**: Where graphs and documentation appear.

**Tip:** Click **File → New File → R Script** to create a script.

---

## Running Basic Math
Run the following in the **Console** (bottom left):

```{r}
2 + 2
5 * 10
100 / 4
```
**Try it!** What happens when you use `-` or `^`?

---

## Creating Variables
Variables store values for later use:

```{r}
x <- 10
y <- 5
sum <- x + y
sum
```

The `<-` (arrow) **assigns** a value to a variable.

---

##  Using Functions
Functions perform actions in R. Try:

```{r}
sqrt(25)    # Square root
round(3.14159, 2)  # Rounding to 2 decimal places
```
**Tip:** Functions use **parentheses `( )`** to take inputs.

---

## Creating a Simple Data Frame (Table)
A **data frame** is like an Excel table:

```{r}
students <- data.frame(
  name = c("Alice", "Bob", "Charlie"),
  age = c(20, 22, 21),
  grade = c(90, 85, 88)
)
students
```
**Try it!** Add another student with a new age and grade.

---

##  Importing Data
To read data from a CSV file:

```{r}
# Uncomment and replace 'yourfile.csv' with the actual file name
# data <- read.csv("yourfile.csv")
# head(data)
```
Use `View(data)` (capital V) to see the dataset in RStudio Cloud.

---

##  Exploring Data
Analyze datasets with:

```{r}
summary(students)   # Summary statistics
names(students)     # Column names
dim(students)       # Number of rows and columns
```
**Try running `str(students)`—what does it show?**

---

##  Making a Quick Plot

```{r}
plot(students$age, students$grade, main="Age vs. Grade", xlab="Age", ylab="Grade")
```
The `$` sign **accesses columns** in a dataset.

---

## Your First R Assignment
Complete these tasks in R:
1. Create two new variables, `a` and `b`, and add them together.
2. Make a data frame with at least **4 rows and 3 columns**.
3. Plot one variable **against another**.

---

## Ready to Explore More?
Try **ggplot2** for advanced visualization:

```{r}
install.packages("ggplot2")
library(ggplot2)
ggplot(students, aes(x=age, y=grade)) + geom_point()
```

Need help? Type `?function_name` in the Console (e.g., `?mean`).


